The binary liquid model for gravity-driven microchannel simulations

### 2010-11-08 Text/paper.tex
Answered Michal's comments. Need to forward it to the supervisor.

\begin{abstract}
The motion towards exploration of nano- and microscale phenomena brings new
challenges to simulation tools. The continuous physics approach is limited on
those scales, and methods such as Molecular Dynamics (MD) or the Lattice
Boltzmann Method (LBM) become more appropriate. This work aims to
reproduce the gravity driven microchannel flow of a free surface liquid with the
help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. The binary
liquid model we use is a continuous-interface multiphase model, which presents
the question of how to properly resolve the interface to obtain proper correlations \todo{between what?}.
In this work we investigate how the binary liquid model parameters and the grid resolution
need to be adjusted to properly simulate microchannel flows.  The presented results
should be of interest for industrial applications and to researchers simulating
Bretherton phenomena.
\end{abstract}

@@

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and the incorporation of
physical phenomena on the microscopic level, as is the case in multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in a continuous way. Such a representation
brings issues with the film thickness resolution versus the interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusion \todo{diffusive?} interface should be resolved in such a way as to have
negligible effect on the physics of the film.

@@

The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is a parameter of the free-energy model  The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
Parameters speficic to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken to depend linearly on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
The surface tension in the framework of the binary liquid model is $\sqrt{8 k A}{9}$ \todo{is this right?  should the 9 be in the
denominator?  is not, perhaps put before the root?}.

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film width.
 \item Spurious currents arising from surface tension discretization
can influence mass transfer in the thin film, thereby comoromising the
overall solution. The bubble should therefore be long enough to decrease the
influence of spurious currents.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is single density, two
viscosities model. The classical Bretherton problem is a free-interface problem
where the film thickness is established by itself. The viscosity ratio should
be large, as indicated in \cite{pagonabarraga-fingers}, however no studies
address this issue. \todo{what's the conclusion of the impact of this issue to our paper then?}
  \item The Bretherton problem addresses bubble flow driven by a pressure
difference, not a body force.  It is however relatively difficult to impose a
pressure difference in the framework of the binary liquid model, as two
simultaneous boundary conditions with the associated gradients and laplacians are required.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is concerned.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability.
 \item Most simulations are done for a circular channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate Laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue.
Our benchmark is therefore done is for an axis-symmetric case. \todo{since this paper
is about a 2d flow, which we already said in the intro, what is the point of discussing
curve boundaries which are only an issue in 3d?}
\end{itemize}
Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow driven by a body force.  An analytical
solution to this problem is to the best of our knowledge not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and the
bubble shape
is different in the front and rear menisci. \citet{wong-films} indicated that the
film width for polygonal capillaries
changes from $\Ca^{1/2}$ on distances $x\propto \Ca^{-1}$ to $\Ca^{2/3}$ for
distances $x \gg \Ca^{-1}$. Therefore, for a bubble to have a formed shape in
the  front meniscus we chose the bubble length to be $5$ channel heights. The
channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions.
Figure \ref{fig:benchmark:sketch} represents a sketch of the geometry
used in the benchmark.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

@@

Here, two different examples are presented that can quantitatively show the
initialization procedure for quite different capillary numbers. In these
examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
That explicitly implies the surface tension to be $\sqrt{\frac{8 k
A}{9}}=0.0377$ and the characteristic length of the interface
$5\sqrt{\frac{k}{A}}=5$ lattice boltzmann units. For the sake of simplicity and
stability the relaxation times for liquid and gas phases are taken as
$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which corresponds
to a gas-liquid viscosity ratio of $10$.
\begin{description}
 \item[I \bm{$Ca=0.005$}]
  The film width for such small Capillary numbers \cite{bretherton} is
given by the equation:
  \begin{equation}
  \frac{h_{\infty}}{H}=1.3375 \Ca^{2/3}=0.03899,
  \end{equation}
  where we assume that $2 H$ or $N_y$ is the channel width in lattice units.

\todo{Therefore one needs to resolve the interface which is around $2$ percent of
the  whole channel width with, for instance $8$ lattice units, where $5$ lattice
units occupies the interface width for given. Note that the interface occupies
$62.5$ percent of the interface. -- these two sentences need to be rewritten for clarity}  Therefore the channel
  width is $400$ lattice units.

  We take the bubble length as $5$ channel widths
  and the distance between bubbles is $2$ bubble lengths. This gives a
grid size of $400 \times 6000$. In terms of computer memory
requirements, the calculations with double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
} \times \,\, 9 \text{ number of
populations}\\
&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with this naive assumption for the $50$ percent resolution of the film
width in comparison with the interface width, one can get the memory
requirement of $345 \mathrm{MB}$.

After resolving the interface one can  calculate the
  velocity from the capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{slug}}}=0.005
\frac{0.0377}{0.66666}=2.82 \cdot10^{-4},
  \end{equation}
  where $\mu_{\mathrm{slug}}=1/3 (2.5-0.5)=0.6666$.
  This velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution: \todo{maybe add a ref, to this eq. from the previous section?}
  \begin{equation}
  \begin{aligned}
  &U_{\mathrm{bubble}}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{N_y^2}{8/3
  (\tau_{\mathrm{liq}}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca
  \end{aligned}
  \end{equation}
  The pressure gradient is then:
  \begin{equation}
  \begin{aligned}
  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
  \end{aligned}
  \end{equation}
  Note that in the case of large grids  height resolution is directly
involved in calculation of the physical time step :since:
\begin{equation}
\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
\end{equation}
where for example the real velocities in the physical world can reach
approximately $1$ m/s and physical channel height up to $1$ mm. In this case one
can easily approximate the physical time step:
\begin{equation}
\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
\end{equation}
which is quote small. A typical simulation needs approximately
$100000-300000$ time steps to come to the steady state solution giving the
physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
probably not enough to find the desired correlations. \todo{correlations between what?}

@@

  The film width for such small Capillary numbers is taken from
  \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
  resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
lattice units are related to the interface resolution.  Therefore the channel
  width is $200$ lattice units. We take the bubble length as $5$ channel widths
  with distance between slugs being $2$ bubble lengths \todo{this I think is the 3rd
  time this particular geometry is mentioned in this paper, perhaps just skip it in this place?}. That gives
the grid size of $200 \times 3000$. The velocity can be
calculated from the Capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
  10^{-3}.
  \end{equation}
  The velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca=3.77\cdot10^{-7}.
  \end{aligned}
  \end{equation}
  Our approach to the initial conditions is  to take values from the previously performed
numerical simulations and to calculate the grid sizes from them, matching the
body force gradient through the solution of Poiseuille flow.

@@

We illustrate the procedure in more detail for the capillary number
$0.05$. As it was pointed out earlier, the film thickness in that case is $6$ percent.
Note that in the case of the half-way bounce-back walls \todo{the bounce-back
proceduce was not described earlier in this paper} which are used in the
simulations one needs to calculate the film thickness as:
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2),
\end{equation}
where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
is related to the effective channel height. Note that it is a big
simplification to impose the boundary in the middle beetween the bounce-back
node and the fluid node which give rise to the viscosity dependent location of
the wall \cite{ginzburg-multireflection}. In the case of $\delta=0.06$ the
initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x-1=15(N_y-2)=1500$.
The bubble is initialized as a rectangular box with coordinates
$y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The force gradient can be estimated through the Poiseuille
profile formula, Eq.~\ref{poiseuille:velocity:center}, and it equals
$1.508 \cdot 10^{-6}$.

>>>>

\begin{abstract}
The motion towards exploration of nano- and microscale phenomena brings new
challenges to simulation tools. The continuous physics approach is limited on
those scales, and methods such as Molecular Dynamics (MD) or the Lattice
Boltzmann Method (LBM) become more appropriate. This work aims to
reproduce the gravity driven microchannel flow of a free surface liquid with the
help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. The binary
liquid model we use is a continuous-interface multiphase model, which presents
the question of how to properly resolve the interface in comparison with the film thickness to
obtain Bretherton correlations.
In this work we investigate how the binary liquid model parameters and the grid resolution
need to be adjusted to properly simulate microchannel flows.  The presented results
should be of interest for industrial applications and to researchers simulating
Bretherton phenomena.
\end{abstract}

@@

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and the incorporation of
physical phenomena on the microscopic level, as is the case in multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in a continuous way. Such a representation
brings issues with the film thickness resolution versus the interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusive interface should be resolved in such a way as to have
negligible effect on the physics of the film.

@@

The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is a parameter of the free-energy model  The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
Parameters speficic to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken to depend linearly on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k A}{9}}$.

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film width.
 \item Spurious currents arising from surface tension discretization
can influence mass transfer in the thin film, thereby comoromising the
overall solution. The bubble should therefore be long enough to decrease the
influence of spurious currents.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is a single density, two
viscosities model. The classical Bretherton problem is a free-interface problem
where the film thickness is established by itself. Thus, the viscosity and density ratios for the
Bretherton problem should
be large. However, in terms of the lattice Boltzmann free-energy framework those issues are not
addressed and will be studied in the current work.
  \item The Bretherton problem addresses bubble flow driven by a pressure
difference, not a body force.  It is however relatively difficult to impose a
pressure difference in the framework of the binary liquid model, as two
simultaneous boundary conditions with the associated gradients and laplacians are required.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is concerned.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability.
 \item Most simulations are done for a circular channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate Laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue.
Our benchmark is therefore done is for a plane-symmetric case in the two-dimensional case. We do
not address the axis-symmetric two-dimensional case since it implies the derivation of additional
mass and force terms for the square based LBM grid \cite{halliday-circular}.
\end{itemize}
Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow driven by a body force.  An analytical
solution to this problem is to the best of our knowledge not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and the
bubble shape
is different in the front and rear menisci. \citet{wong-films} indicated that the
film width for polygonal capillaries
changes from $\Ca^{1/2}$ on distances $x\propto \Ca^{-1}$ to $\Ca^{2/3}$ for
distances $x \gg \Ca^{-1}$. Therefore, for a bubble to have a formed shape in
the  front meniscus we chose the bubble length to be $5$ channel heights. The
channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions.
Figure \ref{fig:benchmark:sketch} represents a sketch of the geometry
used in the benchmark.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

@@

Here, two different examples are presented that can quantitatively show the
initialization procedure for quite different capillary numbers. In these
examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
That explicitly implies the surface tension to be $\sqrt{\frac{8 k
A}{9}}=0.0377$ and the characteristic length of the interface
$5\sqrt{\frac{k}{A}}=5$ lattice boltzmann units. For the sake of simplicity and
stability the relaxation times for liquid and gas phases are taken as
$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which corresponds
to a gas-liquid viscosity ratio of $10$.
\begin{description}
 \item[I \bm{$Ca=0.005$}]
  The film width for such small Capillary numbers \cite{bretherton} is
given by the equation:
  \begin{equation}
  \frac{h_{\infty}}{H}=1.3375 \Ca^{2/3}=0.03899,
  \end{equation}
  where we assume that $2 H$ or $N_y$ is the channel width in lattice units. Thus, the film
thickness occupies around $2$ percent of the whole channel height. Let us impose that $8$ lattice
units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ lattice units
($62.5$ percent) are used to resolve the phase interface. Therefore the channel width is $400$
lattice units.

We take the bubble length as $5$ channel widths
  and the distance between bubbles is $2$ bubble lengths. This gives a
grid size of $400 \times 6000$. In terms of computer memory
requirements, the calculations with double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
} \times \,\, 9 \text{ number of
populations}\\
&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with this naive assumption for the $62.5$ percent resolution of the film
width in comparison with the interface width, one can get the memory
requirement of $345 \mathrm{MB}$.

After resolving the interface one can  calculate the
  velocity from the capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{slug}}}=0.005
\frac{0.0377}{0.66666}=2.82 \cdot10^{-4},
  \end{equation}
  where $\mu_{\mathrm{slug}}=1/3 (2.5-0.5)=0.6666$.
  This velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
  \begin{equation}
  \begin{aligned}
  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
  \end{aligned}
  \end{equation}
  Note that in the case of large grids  height resolution is directly
involved in calculation of the physical time step :since:
\begin{equation}
\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
\end{equation}
where for example the real velocities in the physical world can reach
approximately $1$ m/s and physical channel height up to $1$ mm. In this case one
can easily approximate the physical time step:
\begin{equation}
\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
\end{equation}
which is quote small. A typical simulation needs approximately
$100000-300000$ time steps to come to the steady state solution giving the
physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
probably not enough to find the desired physics correlations taken place on macroscopic time level.

@@

  The film width for such small Capillary numbers is taken from
  \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
  resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
lattice units are related to the interface resolution.  Therefore the channel
  width is $200$ lattice units. Therefore using the assumptions for the benchmark 
the grid size is $200 \times 3000$. The velocity can be
calculated from the Capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
  10^{-3}.
  \end{equation}
  The velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca=3.77\cdot10^{-7}.
  \end{aligned}
  \end{equation}
  Our approach to the initial conditions is  to take values from the previously performed
numerical simulations and to calculate the grid sizes from them, matching the
body force gradient through the solution of Poiseuille flow.

@@

We illustrate the procedure in more detail for the capillary number
$0.05$. As it was pointed out earlier, the film thickness in that case is $6$ percent.
Note that in the case of the half-way bounce-back walls \cite{yu} which are used in the
simulations one needs to calculate the film thickness as:
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2),
\end{equation}
where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
is related to the effective channel height. Note that it is a big
simplification to impose the boundary in the middle beetween the bounce-back
node and the fluid node which give rise to the viscosity dependent location of
the wall \cite{ginzburg-multireflection}. In the case of $\delta=0.06$ the
initial grid size is $N_y=102$, which gives the horizontal grid size as $N_x-1=15(N_y-2)=1500$.
The bubble is initialized as a rectangular box with coordinates
$y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The force gradient can be estimated through the Poiseuille
profile formula, Eq.~\ref{poiseuille:velocity:center}, and it equals
$1.508 \cdot 10^{-6}$.
